---
title: "Anscombe's quartet"
output: html_notebook
---

# Introduction

Anscombe's quartet comprises four data sets that have nearly identical simple
descriptive statistics, yet have very different distributions and appear very
different when graphed. I am going to repeat calculations and graphs using a
`tidyverse` approach.

# Packages

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(kableExtra)
```

# Data

The Anscombe dataset can be found in the base R `datasets` package.

```{r}
anscombe %>% kable() %>% kable_styling()
```

First let's create a `tibble` containing the data.

```{r}
grp1 <- anscombe %>% select(x = x1, y = y1) %>% mutate(group = "Group 1")
grp2 <- anscombe %>% select(x = x2, y = y2) %>% mutate(group = "Group 2")
grp3 <- anscombe %>% select(x = x3, y = y3) %>% mutate(group = "Group 3")
grp4 <- anscombe %>% select(x = x4, y = y4) %>% mutate(group = "Group 4")



data <- bind_rows(grp1, grp2, grp3, grp4) %>%
  
  as_tibble() %>%
  
  mutate(group = as_factor(group))



rm(grp1, grp2, grp3, grp4)



# data %>% kable() %>% kable_styling()
```

# Numerical Summaries

For each of the four datasets let's evaluate some basic numerical summaries.

```{r}
intercept <- function(x, y) {
  model <- lm(y ~ x)
  coefficients(model)[1]
}

slope <- function(x, y) {
  model <- lm(y ~ x)
  coefficients(model)[2]
}

r_squared <- function(x, y){
  model <- lm(y ~ x)
  summary(model)$r.squared 
}

data %>%
  group_by(group) %>%
  summarise(mean_x    = mean(x)         %>% round(1),
            var_x     = var(x)          %>% round(1),
            mean_y    = mean(y)         %>% round(1),
            var_y     = var(y)          %>% round(1),
            corr      = cor(x, y)       %>% round(1),
            intercept = intercept(x, y) %>% round(1),
            slope     = slope(x, y)     %>% round(1),
            r_squared = r_squared(x,y)  %>% round(2)) %>%
  
  kable(col.names = c("Group",
                      "Mean X",
                      "Variance X",
                      "Mean Y",
                      "Variance Y",
                      "Correlation X and Y",
                      "Intercept",
                      "Slope",
                      "R Squared")) %>%
  kable_styling()
```

# Scatterplots

Finally let's plot the four datasets.

```{r plot, fig.align = "center"}
ggplot(data, aes(x = x, y = y)) +
  ggtitle("Anscombe's quartet") +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "red", fullrange = TRUE) +
  facet_wrap(vars(group)) +
  theme_bw(base_size=12) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 24), )
```
# References

This [Wikipedia](https://en.wikipedia.org/wiki/Anscombe%27s_quartet) page and
the references therein. 

