---
title: "Anscombe's Quartet"
format: html
---

# Introduction

Anscombe's quartet comprises four data sets that have nearly identical simple
descriptive statistics, yet have very different distributions and appear very
different when graphed. I am going to repeat calculations and graphs using a
*tidyverse* approach.

# Packages

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)

library(ggthemes)
library(hrbrthemes)

hrbrthemes::import_roboto_condensed()

library(kableExtra)
```

# Data

Load data from csv created in the `make_tidy_csv.R` script.

```{r}
data <- read_csv("anscombe.csv")
```

# Numerical Summaries

For each of the four datasets let's evaluate some basic numerical summaries.

```{r}
intercept <- function(x, y) {
    
    data <- tibble(x = x, y = y)
    
    model <- linear_reg() %>% 
        set_mode(mode = "regression") %>% 
        set_engine(engine = "lm") %>% 
        fit(y ~ x, data = data)
    
    model %>%
        tidy() %>%
        filter(term == "(Intercept)") %>% 
        select(estimate) %>% 
        pull()
}

slope <- function(x, y) {
    
    data <- tibble(x = x, y = y)
    
    model <- linear_reg() %>% 
        set_mode(mode = "regression") %>% 
        set_engine(engine = "lm") %>% 
        fit(y ~ x, data = data)
    
    model %>%
        tidy() %>%
        filter(term == "x") %>% 
        select(estimate) %>% 
        pull()
}

r_squared <- function(x, y) {
    
    data <- tibble(x = x, y = y)
    
    model <- linear_reg() %>% 
        set_mode(mode = "regression") %>% 
        set_engine(engine = "lm") %>% 
        fit(y ~ x, data = data)
    
    glance(model) %>% 
        select(r.squared) %>% 
        pull()
}

data %>%
    
  group_by(group) %>%

  summarise(mean_x    = mean(x)         %>% round(1),
            var_x     = var(x)          %>% round(1),
            mean_y    = mean(y)         %>% round(1),
            var_y     = var(y)          %>% round(1),
            corr      = cor(x, y)       %>% round(2),
            intercept = intercept(x, y) %>% round(2),
            slope     = slope(x, y)     %>% round(2),
            r_squared = r_squared(x,y)  %>% round(2)) %>%
  
  kable(col.names = c("Group",
                      "Mean X",
                      "Variance X",
                      "Mean Y",
                      "Variance Y",
                      "Correlation X and Y",
                      "Intercept",
                      "Slope",
                      "R Squared"),
        
        format = "markdown")
```

# Scatterplots

Finally let's plot the four datasets.

```{r plot, fig.align = "center", message = FALSE, fig.width = 10, fig.height = 10}
plot <- data %>%
    
    ggplot(mapping = aes(x = x, y = y)) +
    
    labs(title = "Anscombe's Quartet") +
    
    geom_point(size = 3, color = "steelblue") +
    
    geom_smooth(method = "lm", se = FALSE, alpha = 0.1,
                
                color = "black", fullrange = TRUE, size = 1.5) +
    
    facet_wrap(~group, labeller = as_labeller(
        c(`Group_1` = "Group 1",
          `Group_2` = "Group 2",
          `Group_3` = "Group 3",
          `Group_4` = "Group 4"))) +
    
    # theme_fivethirtyeight() +
    theme_ipsum() +
    
    theme(strip.text.x = element_text(size = 15))

plot

ggsave("anscombe.png", plot = plot)
```

# References

This [Wikipedia](https://en.wikipedia.org/wiki/Anscombe%27s_quartet) page.