---
title: "Anscombe's Quartet"
format:
    html:
        code-fold: true
---

# Introduction

Anscombe's quartet comprises four data sets that have nearly identical simple
descriptive statistics, yet have very different distributions and appear very
different when graphed. I am going to repeat calculations and graphs using a
*tidy* approach.

# Packages

```{r packages}
#| message: false

library(tidyverse)
library(tidymodels)

library(kableExtra)
```

# Data

Load data from csv created in the `make_tidy_csv.R` script.

```{r}
anscombe_data <- read_csv("anscombe.csv",
                          
                          col_types = list(x = "n", y = "n", group = "f"),
                          
                          show_col_types = FALSE)
```

# Numerical Summaries

For each of the four datasets let's evaluate some basic numerical summaries.

```{r}
intercept <- function(x, y) {
    
    data <- tibble(x = x, y = y)
    
    model <- linear_reg() %>% 
        set_mode(mode = "regression") %>% 
        set_engine(engine = "lm") %>% 
        fit(y ~ x, data = data)
    
    model %>%
        tidy() %>%
        filter(term == "(Intercept)") %>% 
        select(estimate) %>% 
        pull()
}

slope <- function(x, y) {
    
    data <- tibble(x = x, y = y)
    
    model <- linear_reg() %>% 
        set_mode(mode = "regression") %>% 
        set_engine(engine = "lm") %>% 
        fit(y ~ x, data = data)
    
    model %>%
        tidy() %>%
        filter(term == "x") %>% 
        select(estimate) %>% 
        pull()
}

r_squared <- function(x, y) {
    
    data <- tibble(x = x, y = y)
    
    model <- linear_reg() %>% 
        set_mode(mode = "regression") %>% 
        set_engine(engine = "lm") %>% 
        fit(y ~ x, data = data)
    
    glance(model) %>% 
        select(r.squared) %>% 
        pull()
}

anscombe_data %>%
    
    group_by(group) %>%
    
    summarise(mean_x    = mean(x)         %>% round(1),
              var_x     = var(x)          %>% round(1),
              mean_y    = mean(y)         %>% round(1),
              var_y     = var(y)          %>% round(1),
              corr      = cor(x, y)       %>% round(2),
              intercept = intercept(x, y) %>% round(2),
              slope     = slope(x, y)     %>% round(2),
              r_squared = r_squared(x,y)  %>% round(2)) %>%
    
    kable(col.names = c("Group",
                        "Mean X",
                        "Variance X",
                        "Mean Y",
                        "Variance Y",
                        "Correlation X and Y",
                        "Intercept",
                        "Slope",
                        "R Squared"),
          
          format = "markdown")
```

# Scatterplots

Finally let's plot the four datasets.

```{r plot}
(plot <- anscombe_data %>%
     
     ggplot(mapping = aes(x = x, y = y)) +
     
     labs(title = "Anscombe's Quartet") +
     
     geom_smooth(method = "lm", se = FALSE,
                 
                 color = "darkgrey", fullrange = TRUE, size = 0.9) +
     
     geom_point(size = 2.5, color = "deepskyblue") +
     
     coord_fixed(ratio = 0.85) +
     
     facet_wrap(~group, labeller = as_labeller(
         c(`Group_1` = "Group 1",
           `Group_2` = "Group 2",
           `Group_3` = "Group 3",
           `Group_4` = "Group 4"))) +
     
     ggthemes::theme_fivethirtyeight() +
     #hrbrthemes::theme_ipsum() +
     #theme_minimal() +
     #theme(strip.text.x = element_text(size = 15))
     
     theme(axis.title = element_blank(),text = element_text(size = 15)))
```

```{r}
#| echo: false
#| message: false

ggsave("anscombe.png", plot = plot)
```

# References

This [Wikipedia](https://en.wikipedia.org/wiki/Anscombe%27s_quartet) page.