---
title: "Anscombe's Quartet"
output:
    html_document:
        theme: cosmo
---

# Introduction

Anscombe's quartet comprises four data sets that have nearly identical simple
descriptive statistics, yet have very different distributions and appear very
different when graphed. I am going to repeat calculations and graphs using a
*tidy* approach.

# Packages

```{r packages, warning = FALSE, message = FALSE}
library(tidyverse)
#library(tidymodels)

library(kableExtra)

theme_set(theme_minimal(18))
```

# Data

Load data from csv created in the `make_tidy_csv.R` script.

```{r}
data <- read_csv("anscombe.csv")
```

# Numerical Summaries

For each of the four datasets let's evaluate some basic numerical summaries.

```{r}
intercept <- function(x, y) {
  model <- lm(y ~ x)
  coefficients(model)[1]
}

slope <- function(x, y) {
  model <- lm(y ~ x)
  coefficients(model)[2]
}

r_squared <- function(x, y){
  model <- lm(y ~ x)
  summary(model)$r.squared 
}

data %>%
    
  group_by(group) %>%

  summarise(mean_x    = mean(x)         %>% round(1),
            var_x     = var(x)          %>% round(1),
            mean_y    = mean(y)         %>% round(1),
            var_y     = var(y)          %>% round(1),
            corr      = cor(x, y)       %>% round(2),
            intercept = intercept(x, y) %>% round(2),
            slope     = slope(x, y)     %>% round(2),
            r_squared = r_squared(x,y)  %>% round(2)) %>%
  
  kable(col.names = c("Group",
                      "Mean X",
                      "Variance X",
                      "Mean Y",
                      "Variance Y",
                      "Correlation X and Y",
                      "Intercept",
                      "Slope",
                      "R Squared"),
        
        format = "markdown")
```

# Scatterplots

Finally let's plot the four datasets.

```{r plot, fig.align = "center", message = FALSE, fig.width = 10, fig.height = 10}
data %>%
    
    ggplot(mapping = aes(x = x, y = y)) +
    
    labs(title = "Anscombe's Quartet") +
    
    geom_point(size = 3) +
    
    geom_smooth(method = "lm", se = FALSE,
                
                color = "red", fullrange = TRUE, size = 1.5) +
    
    facet_wrap(~group, 
               
               labeller = as_labeller(
                   c(`Group_1` = "Group 1",
                     `Group_2` = "Group 2",
                     `Group_3` = "Group 3",
                     `Group_4` = "Group 4"))
    ) +
    
    theme(plot.title       = element_text(hjust = 0.5, size = 30),
          
          axis.title.x     = element_blank(),
          
          axis.title.y     = element_blank(),
          
          strip.background = element_rect(color="black", size=0.5,
                                          
                                          linetype="solid"),
          
          panel.border     = element_rect(color="black", size=0.5,
                                          
                                          linetype="solid", fill = NA)
    )
```

# References

This [Wikipedia](https://en.wikipedia.org/wiki/Anscombe%27s_quartet) page. 

